---
title: "HW3"
author: "Alisha Wille - afw599"
date: "2025-02-04"
output: pdf_document
---

GitHub link:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
#libraries
library(tidyverse)
library(mosaic)
library(scales)
library(viridis)

# read datasets
gasprices <- read.csv("gasprices.csv")
sclass <- read.csv("sclass.csv")
nbc <- read.csv("nbc_pilotsurvey.csv")
ebay <- read.csv("ebay.csv")

```

## Problem 1

### Theory A

```{r}
# graph for if gas stations charge more if no direct competition is in sight
ggplot(gasprices, aes(x = Price, y = Competitors)) +
  geom_boxplot(fill = "#21908CFF", color = "black") +
  scale_x_continuous(labels = label_dollar()) +
  labs(
    title = "Gas Station Prices if a Competitor is in Sight",
    x = "Price",
    y = "Competitor"
  )

# make table
comp_price <- gasprices |>
  group_by(Competitors = ifelse(Competitors == "Y", "Y", "N")) |>
  summarize(comp_price = round(median(Price), digits = 2)) 

print(comp_price)
```

Theory A states that gas stations charge more if they lack direct competition in sight. According to the data, the median price if there is no competitor in sight was \$1.89, whereas if there was a competitor in sight, the median price was \$1.85. Furthermore, the third quartile of gas prices when there is no competitor in the area is much higher compared to if there is a competitor. Although the difference is only 4 cents by median, the theory is supported by the data.

### Theory B

```{r}
# graph for if the richer the area, the higher the gas prices
ggplot(gasprices, aes(x = Price, y = Income)) +
  geom_point(color = "#21908CFF") +
  geom_smooth(aes(x = Price, y = Income), color = "#FDE725FF") +
  scale_x_continuous(labels = label_dollar()) +
  scale_y_continuous(labels = label_dollar()) +
  labs(
    title = "Gas Station Prices v. Income of Area",
    x = "Price",
    y = "Income"
  )
```

Theory B states that the richer the area, the higher the gas prices. According to the scatterplot, there is a mostly positive correlation between the income of the area and the gas prices. The graph shows a slightly negative correlation until the price hits \$1.80, and shows a positive correlation for all gas prices higher than \$1.80. Therefore, this theory is supported by the data.

### Theory C

```{r}
# graph for if they charge more at a stoplight
ggplot(gasprices, aes(x = Price, y = Stoplight)) +
  geom_boxplot(fill = "#21908CFF", color = "black") +
  scale_x_continuous(labels = label_dollar()) +
  labs(
    title = "Gas Station Prices if by a Stoplight",
    x = "Price",
    y = "Stoplight"
  )

# table
stoplight_price <- gasprices |>
  group_by(Stoplight = ifelse(Stoplight == "Y", "Y", "N")) |>
  summarize(stoplight_price = round(median(Price), digits = 2)) 

print(stoplight_price)
```

Theory C states that gas stations at stoplights charge more. The median gas price for stations not at a stoplight was \$1.84, while the price for stations at a stoplight was \$1.87. However, both the third quartile and maximum value of gas prices for stations not at a stoplight was higher than for the ones that are. Therefore, this theory is not necessarily supported by the data.

### Theory D

```{r}
# if they charge more with direct highwat access
ggplot(gasprices, aes(x = Price, y = Highway)) +
  geom_boxplot(fill = "#21908CFF", color = "black") +
  scale_x_continuous(labels = label_dollar()) +
  labs(
    title = "Gas Station Prices based on Highway Access",
    x = "Price",
    y = "Highway Access"
  )

# table
highway_price <- gasprices |>
  group_by(Highway = ifelse(Highway == "Y", "Y", "N")) |>
  summarize(highway_price = round(median(Price), digits = 2)) 

print(highway_price)
```

Theory D states that highways with direct highway access charge more. The median gas price for stations with direct highway access was \$1.89, while the price for those without direct highway access was \$1.84. Furthermore, the first and third quartiles for stations with direct highway access were much higher than for stations without. Therefore, this theory is supported by the data.

### Theory E

```{r}
# if shell charges more than other brands

# group data by shell v. not
avg_prices <- gasprices |>
  group_by(Brand = ifelse(Brand == "Shell", "Shell", "Other brands")) |>
  summarize(average_price = round(mean(Price), digits = 2)) 

avg_prices_all <- gasprices |>
  group_by(Brand) |>
  summarize(average_price = round(mean(Price), digits = 2)) 

# make graph
ggplot(avg_prices, aes(x = Brand, y = average_price, fill = Brand)) +
  geom_col() +
  scale_fill_manual(values = viridis(2)) +
  geom_text(aes(label = scales::dollar(average_price)), vjust = -0.5, size = 4) + 
  scale_y_continuous(labels = label_dollar()) +
  labs(
    title = "If Shell Charges More for Gas than Other Brands",
    x = "Brand",
    y = "Average Gas Price"
  )

ggplot(avg_prices_all, aes(x = Brand, y = average_price, fill = Brand)) +
  geom_col() +
  scale_fill_manual(values = viridis(4)) + 
  geom_text(aes(label = scales::dollar(average_price)), vjust = -0.5, size = 4) + 
  scale_y_continuous(labels = label_dollar()) +
  labs(
    title = "If Shell Charges More for Gas than Other Brands",
    x = "Brand",
    y = "Average Gas Price"
  )

print(avg_prices)
print(avg_prices_all)
```

Theory E states that Shell charges more for gas than all other non-Shell brands. The average price of gas for Shell was \$1.88, while all other brands had an average of \$1.86. However, when compared to each brand in the dataset, Chevron-Texaco also had an average price of \$1.88 for gas. Therefore, this theory is not supported by the data.

## Problem 2

### Part A

```{r}
# filter dataset
sclass_subset <- sclass |>
  filter(year == 2011 & trim == "63 AMG")

# mileage
sclass_mileage <- sclass_subset$mileage

# define # of bootstrap samples
n_bootstrap_sclass <- 10000

# perform bootstrap
set.seed(123)
bootstrap_sclass_means <- replicate(n_bootstrap_sclass, {
  sample_mileage <- sample(sclass_mileage, size = length(sclass_mileage), replace = TRUE)
  mean(sample_mileage)
})

# confidence interval
sclass_ci_lower <- quantile(bootstrap_sclass_means, 0.025)
sclass_ci_upper <- quantile(bootstrap_sclass_means, 0.975)

# print confidence interval
cat("95% Bootstrap Confidence Interval for Average Mileage:\n")
cat("Lower Bound:", sclass_ci_lower, "\n")
cat("Upper Bound:", sclass_ci_upper, "\n")

# create histogram
sclass_df <- data.frame(means = bootstrap_sclass_means)
ggplot(sclass_df) + 
  geom_histogram(aes(x=bootstrap_sclass_means), fill = "#75D054FF", color = "black") +
   geom_vline(xintercept = sclass_ci_lower, linetype = "dashed", color = "#440154FF", size = 1) +  # Lower CI
  geom_vline(xintercept = sclass_ci_upper, linetype = "dashed", color = "#440154FF", size = 1) +  # Upper CI
   labs(title = "Bootstrap Distribution of Mean Mileage",
       x = "Bootstrapped Mean Mileage",
       y = "Density")

```

A bootstrapped confidence interval for the average mileage of 2011 S-Class 63 AMGs hitting the used-car market was computed. The lower bound of the 95% confidence interval is 26,322.09. The upper bound is 31,862.66.

### Part B

```{r}
# filter dataset
black_subset <- sclass |>
  filter(year == 2014 & trim == "550") |>
  mutate(isBlack = color == "Black")

# make isBlack numeric
black_cars <- as.numeric(black_subset$isBlack)

# set # bootstrap samples
n_bootstrap_black <- 10000

# perform bootstrap
set.seed(456)
bootstrap_black_props <- replicate(n_bootstrap_black, {
  sample_black <- sample(black_cars, size = length(black_cars), replace = TRUE)
  mean(sample_black)  
})

# confidence interval
black_ci_lower <- quantile(bootstrap_black_props, 0.025)  # 2.5th percentile
black_ci_upper <- quantile(bootstrap_black_props, 0.975)  # 97.5th percentile

# print confidence interval
cat("95% Bootstrap Confidence Interval for Proportion of Black Cars:\n")
cat("Lower Bound:", black_ci_lower, "\n")
cat("Upper Bound:", black_ci_upper, "\n")

# create histogram
black_df <- data.frame(means = bootstrap_black_props)
ggplot(black_df) + 
  geom_histogram(aes(x=bootstrap_black_props), fill = "#75D054FF", color = "black") +
   geom_vline(xintercept = black_ci_lower, linetype = "dashed", color = "#440154FF", size = 1) +  # Lower CI
  geom_vline(xintercept = black_ci_upper, linetype = "dashed", color = "#440154FF", size = 1) +  # Upper CI
   labs(title = "Bootstrap Distribution of Black Cars",
       x = "Bootstrapped Black Car Proportions",
       y = "Density")
```

A bootstrapped confidence interval for the proportion of all 2014 S-Class 550s that were painted black was computed. The lower bound of the 95% confidence interval is 0.417, or 41.7%. The upper bound is 0.452, or 45.2%.

## Problem 3

### Part A

```{r}
# filter data
ed_earl_nbc <- nbc |>
  filter(Show == "Living with Ed" | Show == "My Name is Earl")

# separate happy scores by show
ed_happy <- ed_earl_nbc |>
  filter(Show == "Living with Ed") |>
  pull(Q1_Happy)

earl_happy <- ed_earl_nbc |>
  filter(Show == "My Name is Earl") |>
  pull(Q1_Happy)

# compute mean difference
ed_earl_mean_diff <- mean(ed_happy, na.rm = TRUE) - mean(earl_happy, na.rm = TRUE)

# bootstrap samples
set.seed(111)  
ed_earl_bootstrap <- 10000  

ed_earl_diffs <- replicate(ed_earl_bootstrap, {
  ed_sample <- sample(ed_happy, size = length(ed_happy), replace = TRUE)
  earl_sample <- sample(earl_happy, size = length(earl_happy), replace = TRUE)
  mean(ed_sample) - mean(earl_sample)  
})

# confidence interval
ed_earl_ci_lower <- quantile(ed_earl_diffs, 0.025)
ed_earl_ci_upper <- quantile(ed_earl_diffs, 0.975)

# show CI
cat("Living with Ed mean happiness rating:", mean(ed_happy), "\n")
cat("My Name is Earl mean happiness rating:", mean(earl_happy), "\n")
cat("Observed Mean Difference (Ed - Earl):", ed_earl_mean_diff, "\n")
cat("95% Bootstrap Confidence Interval for Mean Difference:\n")
cat("Lower Bound:", ed_earl_ci_lower, "\n")
cat("Upper Bound:", ed_earl_ci_upper, "\n")

# graph
ed_earl_df <- data.frame(differences = ed_earl_diffs)

# Create a histogram with a density curve
ggplot(ed_earl_df, aes(x = differences)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "#24868EFF", color = "black") +
  geom_vline(xintercept = ed_earl_ci_lower, linetype = "dashed", color = "#440154FF", size = 1) +  
  geom_vline(xintercept = ed_earl_ci_upper, linetype = "dashed", color = "#440154FF", size = 1) +
  geom_vline(xintercept = ed_earl_mean_diff, linetype = "solid", color = "#FDE725FF", size = 1) +  
  labs(title = "Bootstrap Distribution of Mean Difference in Happiness Scores",
       x = "Mean Difference (Living with Ed - My Name is Earl)",
       y = "Density")

```

The question I am trying to answer is whether the show "Living with Ed" or "My Name is Earl" produces higher mean happiness scores for viewers. I took the approach of computing the mean happiness rating for each show, then subtracting the ratings to find the difference. I then computed a 95% confidence interval for my findings. I found that "Living with Ed" received a mean happiness rating of 3.927, while "My Name is Earl" received a 3.778. This means that the show "Living with Ed" had a higher score by 0.149. The lower bound of my confidence interval is -0.103, while the upper bound is 0.393. Therefore, I found that the show "Living with Ed" made viewers slightly happier on average due to receiving slightly higher average happiness ratings.

### Part B

```{r}
# filter data
tbl_tala_nbc <- nbc |>
  filter(Show == "The Biggest Loser" | Show == "The Apprentice: Los Angeles")

# separate annoyed scores by show
tbl_annoyed <- tbl_tala_nbc |>
  filter(Show == "The Biggest Loser") |>
  pull(Q1_Annoyed)

tala_annoyed <- tbl_tala_nbc |>
  filter(Show == "The Apprentice: Los Angeles") |>
  pull(Q1_Annoyed)

# compute mean difference
tbl_tala_mean_diff <- mean(tbl_annoyed, na.rm = TRUE) - mean(tala_annoyed, na.rm = TRUE)

# bootstrap samples

tbl_tala_bootstrap <- 10000  

tbl_tala_diffs <- replicate(tbl_tala_bootstrap, {
  tbl_sample <- sample(tbl_annoyed, size = length(tbl_annoyed), replace = TRUE)
  tala_sample <- sample(tala_annoyed, size = length(tala_annoyed), replace = TRUE)
  mean(tbl_sample) - mean(tala_sample)  
})

# confidence interval
tbl_tala_ci_lower <- quantile(tbl_tala_diffs, 0.025)
tbl_tala_ci_upper <- quantile(tbl_tala_diffs, 0.975)

# show CI
cat("The Biggest Loser mean annoyed rating:", mean(tbl_annoyed), "\n")
cat("The Apprentice: Los Angeles mean annoyed rating:", mean(tala_annoyed), "\n")
cat("Observed Mean Difference (TBL - TALA):", tbl_tala_mean_diff, "\n")
cat("95% Bootstrap Confidence Interval for Mean Difference:\n")
cat("Lower Bound:", tbl_tala_ci_lower, "\n")
cat("Upper Bound:", tbl_tala_ci_upper, "\n")

# graph
tbl_tala_df <- data.frame(differences = tbl_tala_diffs)

# Create a histogram with a density curve
ggplot(tbl_tala_df, aes(x = differences)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "#24868EFF", color = "black") +
  geom_vline(xintercept = tbl_tala_ci_lower, linetype = "dashed", color = "#440154FF", size = 1) +  
  geom_vline(xintercept = tbl_tala_ci_upper, linetype = "dashed", color = "#440154FF", size = 1) +
  geom_vline(xintercept = tbl_tala_mean_diff, linetype = "solid", color = "#FDE725FF", size = 1) +  
  labs(title = "Bootstrap Distribution of Mean Difference in Annoyed Scores",
       x = "Mean Difference (The Biggest Loser - The Apprentice: Los Angeles)",
       y = "Density")
```

The question I am trying to answer is whether the show "The Biggest Loser" or "The Apprentice: Los Angeles" produces higher mean annoyed responses for viewers. I took the approach of computing the mean annoyed response for each show, then subtracting the scores to find the difference. I then computed a 95% confidence interval for my findings. I found that "The Biggest Loser" received a mean annoyed response of 2.036, while "The Apprentice: Los Angeles" received a 2.307. This means that the show "The Biggest Loser" had a lower score by -0.271. The lower bound of my confidence interval is -0.527, while the upper bound is -0.018. Therefore, I found that the show "The Biggest Loser" made viewers less annoyed than "The Apprentice: Los Angeles" due to receiving a lower average annoyed response.

### Part C

```{r}
# filter data
dwts_nbc <- nbc |>
  filter(Show == "Dancing with the Stars") |>
  mutate(is_confused = Q2_Confusing >= 4)

# sample proportion
p_hat <- mean(dwts_nbc$is_confused, na.rm = TRUE) # sample prop
n <- sum(dwts_nbc$is_confused) # number valid responses

# standard error for proportion
se <- sqrt((p_hat * (1 - p_hat)) / n)

# confidence interval
z_star <- qnorm(0.975) # z score for 95% CI
dwts_ci_lower <- p_hat - z_star * se
dwts_ci_upper <- p_hat + z_star * se

# results
cat("Estimated Proportion Confused:", round(p_hat, 4), "\n")
cat("95% Confidence Interval for Proportion of Confused Viewers:\n")
cat("Lower Bound:", round(dwts_ci_lower, 4), "\n")
cat("Upper Bound:", round(dwts_ci_upper, 4), "\n")

# graph
dwts_ci_df <- data.frame(
  estimate = p_hat,
  lower = dwts_ci_lower,
  upper = dwts_ci_upper
)

ggplot(dwts_ci_df, aes(x = "Dancing with the Stars", y = estimate)) +
  geom_point(size = 4, color = "#21908CFF") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "#440154FF") +
  labs(
    title = "95% Confidence Interval for Proportion Confused by Dancing with the Stars",
    x = "Show",
    y = "Proportion Confused"
  )

```

The question I am trying to answer is what proportion of American TV watchers would be expected to give a response of 4 or greater to whether "Dancing With the Stars" is confusing. I took the approach of mutating a new variable for reported confused scores of 4 or higher and computing the proportion of this variable to those "not confused". I then computed a 95% confidence interval for my findings. I found that the proportion of confused viewers is 0.077, or 7.7%. The lower bound of my confidence interval is -0.063, or -6.3%, while my upper bound is 0.217, or 21.7%. Therefore, I found that "Dancing With the Stars" is not a confusing show for the large majority of people, due to the low proportion of responses of 4 or higher.

## Problem 4

```{r}
# compute revenue ratio
ebay <- ebay |>
  mutate(revenue_ratio = rev_after / rev_before)

# split treatment and control groups
treatment_group <- ebay |>
  filter(adwords_pause == 1) |>
  pull(revenue_ratio)

control_group <- ebay |>
  filter(adwords_pause == 0) |>
  pull(revenue_ratio)

# compute observed difference in means
ebay_obs_diff <- mean(treatment_group) - mean(control_group)

# bootstrap
set.seed(777)  
n_iterations <- 10000
ebay_bootstrap_diffs <- replicate(n_iterations, {
  ebay_boot_treatment <- sample(treatment_group, size = length(treatment_group), replace = TRUE)
  ebay_boot_control <- sample(control_group, size = length(control_group), replace = TRUE)
  mean(ebay_boot_treatment) - mean(ebay_boot_control)
})

# 95% CI
ebay_ci_lower <- quantile(ebay_bootstrap_diffs, 0.025)
ebay_ci_upper <- quantile(ebay_bootstrap_diffs, 0.975)

# print results
cat("Mean Revenue Ratio for Treatment Group:", mean(treatment_group), "\n")
cat("Mean Revenue Ratio for Control Group:", mean(control_group), "\n")
cat("Observed Difference in Mean Revenue Ratio:", ebay_obs_diff, "\n")
cat("95% Confidence Interval: [", ebay_ci_lower, ",", ebay_ci_upper, "]\n")

# create dataframe
ebay_df <- data.frame(ebay_bootstrap_diffs = ebay_bootstrap_diffs)

# plot bootstrap distribution
ggplot(ebay_df, aes(x = ebay_bootstrap_diffs)) +
  geom_histogram(binwidth = 0.001, color = "#75D054FF", fill = "#75D054FF", alpha = 0.7) +
  geom_vline(xintercept = ebay_ci_lower, color = "#440154FF", linetype = "dashed", size = 1) +
  geom_vline(xintercept = ebay_ci_upper, color = "#440154FF", linetype = "dashed", size = 1) +  
  labs(
    title = "Bootstrap Distribution of Revenue Ratio Differences",
    x = "Difference in Revenue Ratio (Treatment - Control)",
    y = "Frequency"
  )
```

I am trying to answer the question of whether the revenue ratio is the same in the treatment and control designated market areas (DMAs), or if the data favors the idea that paid search advertising on Google creates extra revenue for EBay. For the treatment group, advertising on Google AdWords for the DMA was paused for a month, while advertising on Google AdWords stayed the same in the control group. I took the approach of mutating a new variable in the data for the revenue ratio, and then splitting the data into the treatment and control group. I then computed the mean difference in revenue ratio for the two groups, and computed a 95% confidence interval for my findings. I found that the mean revenue ratio for the treatment group was 0.897, while the mean revenue ratio for the control group was 0.949. The observed mean difference in the groups was -0.052. The lower bound of my 95% confidence interval is -0.091, while the upper bound is -0.013. Therefore, paid search advertising on Google does create extra revenue for EBay due to the revenue ratio being higher in the control group.
